---
# pulled this from Amber's climate report .Rmd for us to modify: https://github.com/nationalparkservice/CCRP-Exposure-Reports/blob/main/CCExposure.Rmd
params:
  name:
    label: Park Name # long name for park?
    value: params$name # park code? not positive how this works
toc: yes
toc_depth: 2
theme: united
output:
  word_document:
    reference_docx: template.docx
bibliography: ExposureReports.bib
csl: ecology.csl
editor_options: 
  markdown: 
    wrap: 79
---

```{r setup, include=FALSE}
# set chunk defaults
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      error = FALSE,
                      message = FALSE)
                      
                      
# read in libraries
source("setup.R")


# set parameters
## path to data folder (from project directory)
path <- "data/"
## park code for this report
park <- "BRCA"


# read in data
load(paste0(path, "/", park, "/", park, "_report_data.RData"))

# match Amber's 'Appendix_Script.R' to create env vars to use for report: https://github.com/nationalparkservice/CCRP-Exposure-Reports/blob/main/Appendix_Script.R
#source("Appendix-Script.R")
```

# Report Title

## Key Findings

High level summary of key findings. Metrics such as projected change in runoff,
etc.

**Recommended Citation**

\newpage

## Project Background

-   Text describing context for project and climate change at parks 

Rising temperatures, changing precipitation regimes, stronger storms, and other
climatic changes are evident across America’s national parks. These changes
have the potential to impact the natural, cultural, and built resources of our
parks, in turn, impacting opportunities to visit and recreate in these spaces.
The National Park Service (NPS) has developed guidance and resources to help
parks incorporate climate considerations into their planning processes
(<https://www.nps.gov/subjects/climatechange/planning.htm>).

The NPS [Planning for a Changing
Climate](https://irma.nps.gov/DataStore/DownloadFile/662814) (@P4CC2021) guide
emphasizes that climate-informed plans should:

1)  Develop forward-looking goals that consider future climatic conditions
    according to the climate projections.
2)  Consider more than one scenario of the future when developing management
    strategies and actions.

Successful climate change planning requires us to understand that we will need
to adapt to some impacts from climate change. Climate change adaptation is
defined as "an adjustment in natural or human systems that moderates harm or
harnesses beneficial opportunities." Timely, forward-looking adaptation can
help conserve cultural and natural resources, develop climate change-ready
infrastructure, safeguard human well-being, and foster a positive visitor
experience.[^1]

[^1]: See [Climate Change
    Glossary](https://irmadev.nps.gov/DataStore/DownloadFile/665664) for full
    definition.

This climate change vulnerability assessment describes historic and projected
changes in hydrology that have potential to impact water supplies at
`r park$UNIT_NAME` (`r params$name`). Future hydrology is derived from climate
and water balance models that explain different ways that the future might
evolve at `r params$name` due to climate change. These projections help parks
make short- and long-term decisions that avoid surprises and costly mistakes.
The approach of planning around multiple projections is a practical response to
the impossibility of precisely predicting greenhouse gas emissions and how our
climate will respond to them. [Runyon et al.
(2024)](https://www.nps.gov/subjects/climatechange/climatefutures.htm) and [XX
citation to wbm pub] (link to wbm pub) includes an in-depth description of
methods used for these analyses.

-   Regional specific expected changes

Several climate-related changes to hydrology could impact water supplies at
Inter-mountain region (IMR) National Parks.

-   Reduced snowpack and earlier melt: Rising temperatures will lead to a
    decrease in winter snowfall and an earlier spring melt. This will result in
    less water being stored in snowpack, which is a critical source for
    replenishing rivers and streams throughout the year.

-   Increased droughts and water scarcity: With less snowpack and altered
    precipitation patterns, droughts are expected to become more frequent and
    severe. This will lead to decreased streamflow, lower reservoir levels, and
    potential water shortages for agriculture, municipalities, and ecosystems.

-   Shifts in Streamflow Timing: Earlier snowmelt will cause peak streamflow to
    occur earlier in the spring, potentially leading to increased likelihood of
    flooding during that time. However, with less overall water stored as snow,
    streamflow will likely be lower during the summer months when water demands
    are typically highest.

-   Changes in Precipitation Patterns: While overall precipitation may not
    change dramatically, the pattern of precipitation is expected to shift.
    There might be an increase in extreme precipitation events (heavy rain and
    snowfall) interspersed with longer dry periods. This can lead to flash
    flooding and soil erosion during heavy rain events, while longer dry
    periods exacerbate drought conditions.

-   Potential Impacts on Water Quality: Lower streamflow can lead to increased
    water temperature and reduced dilution of pollutants. This can have
    negative consequences for aquatic ecosystems and water quality for human
    consumption. "

```{r background_fig}


```

# Water budget

Precipitation that falls to the ground can follow various paths: - Some
infiltrates the soil, recharging groundwater and potentially sustaining plant
life. - Some evaporates back into the atmosphere, especially from surfaces like
leaves and open water. - A small portion might be stored as surface water in
lakes, ponds, or wetlands. - The rest flows across the land surface as runoff,
eventually feeding streams, rivers, and ultimately the ocean.

Each path represents a component of the "water budget". A water budget is like
a financial budget for water. It tracks all the water that enters
(precipitation), exits (evapotranspiration, runoff), and gets stored
(groundwater, soil moisture) in the system. Components of the water budget like
recharge and runoff represent annual water inputs for the vast majority of park
water systems (i.e., wells and diversions). By evaluating historical and
projected changes in a water budget over time, we can gain valuable insights
about future water availability and potential stresses.

```{r background_fig}


```


# Demand
