---
title: "Explore WBM Data Pull"
author: "Caitlin Mothes, PhD"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("setup.R")

# load in report data to get BRCA supply and boundary to test pulling
load("data/park/BRCA/BRCA_report_data.RData")

# get singl water supply point to test with
test_point <- POD_supply[1,]

# need to increase timeout option for downloading large daily files
options(timeout=10000)
```


## Explore Pulling Future WBM (edits to `getFutureWBM.R`)


### Create new function to pull future WBM data
```{r}
get_future_wbm <- function(park, aoi, years, macas, wb_var, path = "data/park/") {
  

  # check that path to save files to exists
  if (!dir.exists(paste0(getwd(), "/", path, "/", park, "/wbm_daily_rasters/"))) {
    dir.create(file.path(paste0(getwd(), "/", path, "/", park, "/wbm_daily_rasters/")), showWarnings = FALSE)
  }
  
  # create df of all variable, year, scenario combinations
  map_dat <- tidyr::crossing(years, wb_var, macas) %>%
    tidyr::separate_wider_delim(macas, delim = ".", names = c("maca", "rcp"))
  
  
  # transform aoi to crs of WBM data
  aoi <- st_transform(aoi, crs = "+proj=lcc +lat_0=42.5 +lon_0=-100 +lat_1=25 +lat_2=60 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs")
  
  # download each raster file to temp directory
  
  ## test using furrr
  # plan
  #plan(multisession, workers = parallelly::availableCores() - 1)
  
  rasters <- future_map(1:nrow(map_dat), function(x) {
    call <- paste0(
      "/vsicurl/",
      "http://www.yellowstone.solutions/thredds/fileServer/daily_or_monthly/gcm/",
      map_dat$rcp[x],
      "/",
      map_dat$maca[x],
      "/V_1_5_",
      map_dat$years[x],
      "_",
      map_dat$maca[x],
      "_",
      map_dat$rcp[x],
      "_",
      map_dat$wb_var[x],
      ".nc4"
    )
    
    r <- terra::rast(call)
    
    #  ## saved CRS and extent from raw downloaded file, that info does not attached to files pulled directly from URL
    # crs(r) <- "+proj=lcc +lat_0=42.5 +lon_0=-100 +lat_1=25 +lat_2=60 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"
    # 
    # ext(r) <- c(-2060750, 2639250, -1915500, 1384500)
    # 
    # # crop to aoi
    # r <- terra::crop(r, aoi)
    
    # assign informative names to each raster
    names(r) <- paste(map_dat$maca[x], map_dat$years[x], names(r), sep = "_")
  
    return(r)
    
  })
  
  ## combine all to single stack
  final_stack <- terra::rast(rasters)
  
  crs(final_stack) <- "+proj=lcc +lat_0=42.5 +lon_0=-100 +lat_1=25 +lat_2=60 +x_0=0 +y_0=0 +ellps=WGS84 +units=m +no_defs"
  
  ext(final_stack) <- c(-2060750, 2639250, -1915500, 1384500)
  
  # crop to aoi
  final_stack <- terra::crop(final_stack, aoi)
  
  return(final_stack)
  # save final raster stack to file
  # terra::writeRaster(
  #   final_stack,
  #   filename = paste0(
  #     getwd(),
  #     "/",
  #     path,
  #     "/",
  #     park,
  #     "/wbm_daily_rasters/",
  #     park,
  #     "_",
  #     wb_var,
  #     "_selected_cfs_",
  #     years[1],
  #     "_",
  #     years[length(years)],
  #     ".tif"
  #   ),
  #   overwrite = TRUE
  # )
  
}
```


### Test function

Use `profvis` to check for any bottle necks
```{r}
library(profvis)

start <- Sys.time()

profvis({get_future_wbm(
  park = park,
  aoi = st_buffer(park_boundary, 20000),
  years = 2023,
  wb_var = "runoff",
  macas = select_cfs$GCM
)
})

end <- Sys.time()

end - start


```

Read in the raster output to make sure it looks right
```{r}
daily_runoff <- rast("data/park/BRCA/wbm_daily_rasters/BRCA_runoff_selected_cfs_2023_2070.tif")
```


# Run function for full park data
Testing out adding `future_map()` to the function cut it down by 1 minute.

```{r}
plan(multisession, workers = parallelly::availableCores() - 1)

start <- Sys.time()

#walk(
test <- future_map(
  c(#"soil_water",
    "runoff"),
    #"rain",
    #"accumswe",
    #"PET",
    #"Deficit",
    #"AET"),
    ~ get_future_wbm(
      park = park,
      aoi = st_buffer(park_boundary, 20000),
      years = 2023:2070,
      wb_var = .x,
      macas = select_cfs$GCM
    )
  )

  
end <- Sys.time()

end - start
```

# Parallelize it with `furrr`
Cut the time down to 32 minutes (from ~70)

```{r}
library(furrr)
library(tictoc)

# plan
plan(multisession, workers = parallelly::availableCores() - 1)

tic()

future_walk(
      c(
        "soil_water",
        "runoff",
        "rain",
        "accumswe",
        "PET",
        "Deficit",
        "AET"
      ),
      ~ get_wbm_grids(
        park = park,
        aoi_name = "test_furrr_2",
        aoi = watersupply_watershed,
        wb_var = .x,
        macas = select_cfs$GCM,
        path = path
      ),
      .progress = TRUE
    )

toc()

# sets processing back to default
plan(sequential)

```

# Cleaning outputs for `get_wbm_aoi`

Test use of `dtplyr` over `dplyr`

```{r}
# list of rasters
wbm_aoi <- list.files("data/park/BRCA/wbm_daily_rasters/",full.names = TRUE) %>% 
            grep("watersupply_watershed", ., value=TRUE) 


```

rast_read function:

Version 1 (dplyr) took 15 seconds for one raster. Here trying out `dtplyr`

```{r}
 rast_read <- function(rast_list) {
    
   terra::rast(rast_list) %>%
     terra::as.data.frame(xy = TRUE) %>%
     data.table() %>% 
     #dtplyr::lazy_dt() %>%
     # remove column names that have NA for the data (this removes the -32767 vals)
     dplyr::select(-contains("_NA")) %>%
     pivot_longer(-(x:y), names_to = "var", values_to = "val") %>%
     # Add historic to name
     dplyr::mutate(var =
                     if_else(
                       grepl(
                         '^accumswe|^runoff|^rain|^PET|^AET|^soil_water|^Deficit',
                         var
                       ),
                       paste0("gridmet.historical_", var),
                       var
                     ),
                   val = val / (25.4 * 10)) %>% # 10 = conversion factor, 25.4 = mm to in
     dplyr::mutate(var = gsub("soil_water", "soilwater", var)) %>%
      # KEC - separate is VERY slow - can anyone suggest something quicker?
      separate(var, 
               into = c("gcm", "wbm_var", "date"), 
               sep = "_") %>%
      dplyr::mutate(wbm_var = 
               str_replace_all(
                 wbm_var, 
                 c("runoff" = "runoff_in", 
                   "rain" = "rain_in", 
                   "AET" = "aet_in",
                   "PET" = "pet_in",
                   "accumswe" = "accumswe_in",
                   "soilwater" = "soil_water_in",
                   "Deficit" = "deficit_in")),
             date = as.Date(date))
  }
```

Run the function
```{r}
system.time(test <- rast_read(wbm_aoi[1]))

```

Cut time down from 15 seconds to 3 seconds per file, but brining df back into env takes a while...

Run for all files:
```{r}
 wbm_aoi_cleaned <- wbm_aoi %>%
    purrr::map_dfr(~ rast_read(.)) %>%
    dplyr::select(-c(x,y)) %>%
    group_by(date, wbm_var, gcm) %>%
    summarize_all(.funs = 
                    list(val = ~ mean(val, na.rm = TRUE)), 
                  .groups = "keep") %>%
    pivot_wider(names_from = "wbm_var", values_from = "val") 
```


# Compare dtplyr and data.table()

Dtplyr = 105 seconds
```{r}
plan(multisession)

rast_read <- function(rast_list) {
    
   terra::rast(rast_list) %>%
     terra::as.data.frame(xy = TRUE) %>%
     #data.table() %>% 
     dtplyr::lazy_dt() %>%
     # remove column names that have NA for the data (this removes the -32767 vals)
     dplyr::select(-contains("_NA")) %>%
     pivot_longer(-(x:y), names_to = "var", values_to = "val") %>%
     # Add historic to name
     dplyr::mutate(var =
                     if_else(
                       grepl(
                         '^accumswe|^runoff|^rain|^PET|^AET|^soil_water|^Deficit',
                         var
                       ),
                       paste0("gridmet.historical_", var),
                       var
                     ),
                   val = val / (25.4 * 10)) %>% # 10 = conversion factor, 25.4 = mm to in
     dplyr::mutate(var = gsub("soil_water", "soilwater", var)) %>%
      # KEC - separate is VERY slow - can anyone suggest something quicker?
      separate(var, 
               into = c("gcm", "wbm_var", "date"), 
               sep = "_") %>%
      dplyr::mutate(wbm_var = 
               str_replace_all(
                 wbm_var, 
                 c("runoff" = "runoff_in", 
                   "rain" = "rain_in", 
                   "AET" = "aet_in",
                   "PET" = "pet_in",
                   "accumswe" = "accumswe_in",
                   "soilwater" = "soil_water_in",
                   "Deficit" = "deficit_in")),
             date = as.Date(date)) %>% 
    as_tibble()
}

system.time(test1 <- wbm_aoi %>%
    furrr::future_map_dfr(~ rast_read(.x)))

```

Data.table = 47 seconds
```{r}
plan(multisession)

rast_read <- function(rast_list) {
    
   terra::rast(rast_list) %>%
     terra::as.data.frame(xy = TRUE) %>%
     data.table() %>% 
     #dtplyr::lazy_dt() %>%
     # remove column names that have NA for the data (this removes the -32767 vals)
     dplyr::select(-contains("_NA")) %>%
     pivot_longer(-(x:y), names_to = "var", values_to = "val") %>%
     # Add historic to name
     dplyr::mutate(var =
                     if_else(
                       grepl(
                         '^accumswe|^runoff|^rain|^PET|^AET|^soil_water|^Deficit',
                         var
                       ),
                       paste0("gridmet.historical_", var),
                       var
                     ),
                   val = val / (25.4 * 10)) %>% # 10 = conversion factor, 25.4 = mm to in
     dplyr::mutate(var = gsub("soil_water", "soilwater", var)) %>%
      # KEC - separate is VERY slow - can anyone suggest something quicker?
      separate(var, 
               into = c("gcm", "wbm_var", "date"), 
               sep = "_") %>%
      dplyr::mutate(wbm_var = 
               str_replace_all(
                 wbm_var, 
                 c("runoff" = "runoff_in", 
                   "rain" = "rain_in", 
                   "AET" = "aet_in",
                   "PET" = "pet_in",
                   "accumswe" = "accumswe_in",
                   "soilwater" = "soil_water_in",
                   "Deficit" = "deficit_in")),
             date = as.Date(date))
}

system.time(test2 <- wbm_aoi %>%
    furrr::future_map_dfr(~ rast_read(.x)))
```


# Test final wbm_aoi function

```{r}
test <- get_wbm_aoi(park, aoi = watersupply_watershed, aoi_name = "watersupply_watershed", gcm = select_cfs$GCM, aoi_average = FALSE, download = FALSE)
```



```{r}
test_average <- get_wbm_aoi(park, aoi = watersupply_watershed, aoi_name = "watersupply_watershed", gcm = select_cfs$GCM, aoi_average = TRUE, download = FALSE)
```

